Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-3\baselineskip}

### 2017-03-21 paper/ms.tex
adding HODLR benchmark plots

Another commonly used restrictive method uses the fact that, in the limit of
evenly-spaced data and homoscedastic uncertainties, the covariance matrix is
``Toeplitz'' \citep[for example][]{Dillon:2013}.
There are exact methods for solving Toeplitz matrix equations that scale as
$\mathcal{O}(N\,\log N)$ and methods for computing determinants exactly in
$\mathcal{O}(N^2)$ or approximately in $\mathcal{O}(N\,\log N$
\citep{Wilson:2014}.
The Toeplitz method is, in some ways, more flexible than \celerite\ because it
can be used with any stationary kernel but it requires uniformly spaced data
and the scaling is worse than \celerite\ so it will, in general, be less
efficient when applied to large datasets.

@@

@@

Using \eq{celerite-psd}, we find that for a single \celerite\ term, this
requirement is met when
\begin{eqnarray}\eqlabel{psd-req}
\frac{(a_j\,c_j+b_j\,d_j)\,({c_j}^2+{d_j}^2)+(a_j\,c_j-b_j\,d_j)\,\omega^2}
{\omega^4+2\,({c_j}^2-{d_j}^2)\,\omega^2+{({c_j}^2+{d_j}^2)}^2} > 0 \quad.
\end{eqnarray}
The denominator is positive for all $c_j \ne 0$ and it can be shown that, when
$c_j=0$, \eq{psd-req} is satisfied for all $\omega \ne d_j$ where the power is
identically zero.
Therefore, when $c_j \ne 0$, we require that the numerator is positive for all
$\omega$.
This requirement can also be written as
\begin{eqnarray}
a_j\,c_j > -b_j\,d_j \\
a_j\,c_j > b_j\,d_j \quad.
\end{eqnarray}
Furthermore, we can see that $a_j$ must be positive since $k(0) = a_j$ should
should be positive and, similarly, by requiring the covariance to be finite at
infinite lag, we obtain the constraint $c_j \ge 0$.
Without loss of generality, we can choose $d_j \ge 0$ and find the final
constraint:
\begin{eqnarray}
    -|b_j\,d_j| < a_j\,c_j < |b_j\,d_j|
\end{eqnarray}

>>>>

Another commonly used restrictive method uses the fact that, in the limit of
evenly-spaced data and homoscedastic uncertainties, the covariance matrix is
``Toeplitz'' \citep[for example][]{Dillon:2013}.
There are exact methods for solving Toeplitz matrix equations that scale as
$\mathcal{O}(N\,\log N)$ and methods for computing determinants exactly in
$\mathcal{O}(N^2)$ or approximately in $\mathcal{O}(N\,\log N)$
\citep{Wilson:2014}.
The Toeplitz method is, in some ways, more flexible than \celerite\ because it
can be used with any stationary kernel but it requires uniformly spaced data
and the scaling is worse than \celerite\ so it will, in general, be less
efficient when applied to large datasets.

@@

Another GP method that has been used extensively in astronomy is the
hierarchical off-diagonal low rank \citep[HODLR,][]{Ambikasaran:2016} solver.
This method exploits the fact that many commonly used kernel functions produce
``smooth'' matrices to approximately compute the GP likelihood with the
scaling $\mathcal{O}(N\,\log^2 N)$.
This method has the advantage that, unlike \celerite, it can be used with any
kernel function but, in practice, the cost can still prove to be prohibitively
high for multi-dimensional inputs.

Many other approximate methods for scaling GP inference exist and we will make
no attempt to discuss the comparisons exhaustively.

@@

Using \eq{celerite-psd}, we find that for a single \celerite\ term, this
requirement is met when
\begin{eqnarray}\eqlabel{psd-req}
\frac{(a_j\,c_j+b_j\,d_j)\,({c_j}^2+{d_j}^2)+(a_j\,c_j-b_j\,d_j)\,\omega^2}
{\omega^4+2\,({c_j}^2-{d_j}^2)\,\omega^2+{({c_j}^2+{d_j}^2)}^2} > 0 \quad.
\end{eqnarray}
The denominator is positive for all $c_j \ne 0$ and it can be shown that, when
$c_j=0$, \eq{psd-req} is satisfied for all $\omega \ne d_j$ where the power is
identically zero.
Therefore, when $c_j \ne 0$, we require that the numerator is positive for all
$\omega$.
This requirement can also be written as
\begin{eqnarray}
a_j\,c_j > -b_j\,d_j \\
a_j\,c_j > b_j\,d_j \quad.
\end{eqnarray}
Furthermore, we can see that $a_j$ must be positive since $k(0) = a_j$ should
should be positive and, similarly, by requiring the covariance to be finite at
infinite lag, we obtain the constraint $c_j \ge 0$.
Combining these results, we find the constraint
\begin{eqnarray}
    -|b_j\,d_j| < a_j\,c_j < |b_j\,d_j| \quad.
\end{eqnarray}
